### 2013-05-15 docs/Jesse/Dissertation/problemRange.tex

We consider the transient compressible Navier-Stokes equations. For simplicity, we present them in two spatial dimensions. Each equation of the Navier-Stokes system represents the conservation of some physical quantity in the behavior of a fluid inside a general control volume.\footnote{The derivation of the compressible Navier-Stokes equations is a standard result of the Reynolds transport theorem, and can be found in many elementary fluid dynamics books.}

\begin{itemize}
\item{\textbf{Mass conservation}}
\begin{align*}
\pd{\rho}{t} + \div \vecttwo{\rho u }{\rho v} &= 0
\end{align*}

@@

We assume our fluid satisfies standard stress laws for $\boldsymbol \sigma$ and $\boldsymbol q$ as well. For viscous stresses $\boldsymbol \sigma$, we assume a Newtonian fluid
\begin{align*}
\sigma_{ij} &= \mu(u_{i,j}+u_{j,i}) + \lambda u_{k,k}\delta_{ij}.
\end{align*}
The coefficients $\lambda$ and $\mu$ are known as the viscosity and bulk viscosity, respectively. The bulk viscosity is often set implicitly through $2\mu + 3\lambda = 0$, known as Stokes' hypothesis. However, since the effect of bulk viscosity can become important for compressible flows, we treat both coefficients separately. In general, $\mu$ and $\lambda$ are functions of temperature, obeying the power law
\[
\mu = \left(\frac{T}{T_0}\right)^\beta,
\]
where $T_0$ is a reference temperature. We choose $\beta = 2/3$ in this case.

We assume our fluid satisfies Fourier's law, which relates the heat flux $\boldsymbol q$ to the gradient of the temperature through
\begin{align*}
{\boldsymbol q} &= \kappa \grad T,
\end{align*}
where $\kappa$ is generally a function of temperature.

@@

Under appropriate assumptions on the behavior of density and temperature, the behavior of the compressible Navier-Stokes equations can be sufficiently represented by the incompressible Navier-Stokes equations for low speed laminar fluid flows. The study of the incompressible Navier-Stokes equations is an open area in mathematics, and is one of the most famous Millenium Problems posed by the Clay Mathematics Institute. The equations of incompressible flow pose a difficult problem computationally as well, in part due to the problem of the simulation of turbulent phenomena.

@@

The linearized Navier-Stokes equations are the result of small perturbation assumptions applied to the full Navier-Stokes equations. Under such assumptions, the flow is largely uniform, with slight variations that are small compared to the uniform stream velocity. We are interested in the linearized Navier-Stokes equations mainly for mathematical purposes - as the solution to the full Navier-Stokes equations involves a series of solutions for linearized Navier-Stokes, we wish to investigate the behavior of our numerical method with respect to this system.

@@

The scalar convection-diffusion equation models mathematically the distribution of the concentration $u$ of a substance in a medium due to both convective and diffusive effects. The general form of the equation in $\mathbb{R}^n$ is 
\[
\beta\cdot \grad u - \epsilon \Delta u = f,
\]
where $\beta \in \mathbb{R}^n$ is a vector specifying the direction of convection.

Scalar convection-diffusion has significant historical importance, as it is the prototypical model problem for solving the full Navier-Stokes equations --- most stabilized methods consider first the scalar convection-diffusion equation as a test case before attempting a solution of the full Navier-Stokes equations. As discussed previously, an important feature of the convection-diffusion equation is that solutions can develop boundary layers whose thickness depends on the viscosity, a physical feature found in most applications of interest for compressible flow.

@@

The pure convection equation is a result of neglecting the viscous term in the convection-diffusion equation. Physically speaking, these assumptions correspond to the inviscid limit, as well as a particular class of boundary conditions (for example, the wall boundary condition $u = 0$ and prescribed inflow condition may be incompatible in the inviscid limit).  The Euler equations are likewise a result of neglecting the viscous terms in the Navier-Stokes equations.  However, these problems can be ill-posed in the continuous setting.  Take, for example, the vortex problem in Figure~\ref{fig:convCirc}.  A feature of the convection equation is that there is no crosswind diffusion - thus, materials do not mix across streamlines.  However, for the vortex problem, this also implies that the solution on any closed streamline can take any arbitrary value, and is thus undefined.

\begin{figure}[!h]
\centering
\includegraphics[scale = .22]{figs/convCirc.png}
\caption{Setup for the vortex problem.}
\label{fig:convCirc}
\end{figure}
Formally speaking, the solution to the vortex problem is taken to be the solution to the convection-diffusion equation (with appropriate outflow boundary conditions) as the viscosity tends towards zero, in which case, the solution in the interior would be uniformly zero.  This motivates the need for \emph{artificial viscosity} methods with which to regularize inviscid solutions.  The topic is expansive, and we direct the reader towards \cite{Barter} for a more detailed discussion of past and present artificial viscosity methods.

There is still much interest in solving the inviscid Euler equations in industry --- the full Navier-Stokes models often prove difficult to solve due to the mathematical nature of the equations, and the coupling of the Euler equations with boundary layer models has been successful in simulating many phenomena in compressible flow \cite{BoeingDrela}.

>>>>

We consider the transient compressible Navier-Stokes equations. For simplicity, we present them in two spatial dimensions. Each equation of the Navier-Stokes system represents the conservation of some physical quantity in the behavior of a fluid inside a general control volume.\footnote{The derivation of the compressible Navier-Stokes equations is a standard result of the Reynolds transport theorem, and can be found in many elementary fluid dynamics books. See \cite{Emanuel:994127} for one example.}

In 2D, the classical form of the Navier-Stokes equations involve the fluid density $\rho$, velocity in the $x$ and $y$ directions $u$ and $v$ (or $u_1$ and $u_2$), respectively, temperature $T$, energy per unit mass $e$, and stress and heat flux vectors $\boldsymbol \sigma_i$ and $\vec{q}$. The equations are as follows:
\begin{itemize}
\item{\textbf{Mass conservation}}
\begin{align*}
\pd{\rho}{t} + \div \vecttwo{\rho u }{\rho v} &= 0
\end{align*}

@@

We assume our fluid satisfies standard stress laws for $\boldsymbol \sigma$ and $\boldsymbol q$ as well. For viscous stresses $\boldsymbol \sigma$, we assume a Newtonian fluid
\begin{align*}
\sigma_{ij} &= \mu(u_{i,j}+u_{j,i}) + \lambda u_{k,k}\delta_{ij}.
\end{align*}
The coefficients $\lambda$ and $\mu$ are the viscosity and bulk viscosity, respectively. The bulk viscosity is often set implicitly through $2\mu + 3\lambda = 0$, known as Stokes' hypothesis. However, since the effect of bulk viscosity can become important for compressible flows, we treat both coefficients separately. In general, $\mu$ and $\lambda$ are functions of temperature, obeying the power law
\[
\mu = \left(\frac{T}{T_0}\right)^\beta,
\]
where $T_0$ is a reference temperature. We choose $\beta = 2/3$ in this case.

We assume our fluid satisfies Fourier's law, which relates the heat flux $\boldsymbol q$ to the gradient of the temperature through
\begin{align*}
{\boldsymbol q} &= \kappa \grad T,
\end{align*}
where $\kappa$, the coefficient of heat conductivity, is generally a function of temperature.

@@

Under appropriate assumptions on the behavior of density and temperature, the behavior of the compressible Navier-Stokes equations can be sufficiently represented by the incompressible Navier-Stokes equations for some fluid flows.  For example, the incompressible Navier-Stokes equations accurately model nearly incompressible mediums such as water, as well as low Mach number flows of compressible fluids. The study of the incompressible Navier-Stokes equations is an open area in mathematics, and is one of the most famous Millenium Problems posed by the Clay Mathematics Institute. The equations of incompressible flow pose a difficult problem computationally as well, in part due to the problem of the simulation of turbulent phenomena.

@@

The linearized Navier-Stokes equations are the result of small perturbation assumptions applied to the full Navier-Stokes equations. Under such assumptions, the flow in a domain consists only of slight variations (to a given background flow) that are small compared to the magnitude of the free stream velocity.  Mathematically speaking, the linearized Navier-Stokes equations are the results of the linearization of the full equations with respect to a specific background flow.  The linearized NS equations are frequently referred to as an ``acoustic problem'', as they are used to study sound propagation.

We are interested in the linearized Navier-Stokes equations mainly for mathematical purposes - as the solution to the full Navier-Stokes equations involves a series of solutions for linearized Navier-Stokes, we wish to investigate the behavior of our numerical method with respect to this system.

@@

Recall that the scalar convection-diffusion equation models mathematically the distribution of the concentration $u$ of a substance in a medium due to both convective and diffusive effects. 
%The general form of the equation in $\mathbb{R}^n$ is 
%\[
%\beta\cdot \grad u - \epsilon \Delta u = f,
%\]
%where $\beta \in \mathbb{R}^n$ is a vector specifying the direction of convection.
Scalar convection-diffusion has significant historical importance, as it is the prototypical model problem for solving the full Navier-Stokes equations --- most stabilized methods consider first the scalar convection-diffusion equation as a test case before attempting a solution of the full Navier-Stokes equations. As discussed previously, an important feature of the convection-diffusion equation is that solutions can develop boundary layers whose thickness depends on the viscosity, a physical feature found in most applications of interest for compressible flow.

@@

The pure convection equation is a result of neglecting the viscous term in the convection-diffusion equation. Physically speaking, these assumptions correspond to the inviscid limit, as well as a particular class of boundary conditions (for example, a prescribed inflow condition may be incompatible with the wall boundary condition $u = 0$ in the inviscid limit).  The Euler equations are likewise a result of neglecting the viscous terms in the Navier-Stokes equations.  However, these problems can be ill-posed in the continuous setting.  Take, for example, the vortex problem in Figure~\ref{fig:convCirc}.  A feature of the convection equation is that there is no crosswind diffusion - thus, materials do not mix across streamlines.  However, for the vortex problem, this also implies that the solution on any closed streamline can take any arbitrary value, and is thus undefined.

\begin{figure}[!h]
\centering
\includegraphics[scale = .22]{figs/convCirc.png}
\caption{Setup for the vortex problem.}
\label{fig:convCirc}
\end{figure}
Formally speaking, the solution to the vortex problem is taken to be the solution to the convection-diffusion equation (with appropriate outflow boundary conditions) as the viscosity tends towards zero, in which case, the solution in the interior would be uniformly zero (this technique is referred to in mathematical literature as the ``vanishing viscosity" method, and is used to define unique solutions in the inviscid limit).  This motivates the need for \emph{artificial viscosity} methods with which to regularize inviscid solutions.  The topic is expansive, and we direct the reader towards \cite{Barter} for a more detailed discussion of past and present artificial viscosity methods.

The full Navier-Stokes models have proven difficult to solve due to the mathematical nature of the equations --- due to the lack of robustness of most methods, solving the Navier-Stokes for high Reynolds numbers requires very fine meshes and is an incredibly expensive task.  Additionally, the problem of turbulence for high Reynolds numbers further complicates the Navier-Stokes solutions for high speed compressible flow.  Without turbulence models, turbulent effects can prevent convergence to a solution.  However, common turbulence models, such as Reynolds Averaged Navier-Stokes (RANS), can lead to nonphysical solutions, such as the existence of a steady-state solution when there is none.

In comparison, the coupling of the inviscid Euler equations with boundary layer models has been successful in simulating many phenomena in compressible flow at a computational cost orders of magnitude below that of the full Navier-Stokes equations \cite{BoeingDrela}.  The method has been extended to a wide array of physical conditions, and is an active area of current research in both industry and academia.
